---
title: "Chapter3 Exercises"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

3.2.4 Exercises
1) Run ggplot(data = mpg). What do you see?
```{r}
ggplot(data = mpg)
```
A blank white screen.

2)How many rows are in mpg? How many columns?
```{r}
mpg
```
11 columns.

3) What does the drv variable describe? Read the help for ?mpg to find out.
```{r}
?mpg
```
drv describes f = front-wheeldrive, r = rear wheel drive, 4 = 4wd

4)Make a scatterplot of hwy vs cyl.
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = hwy, y = cyl))
```

5) What happens if you make a scatterplot of class vs drv? Why is the plot not useful?
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = drv))
```
The reason it's not helpful is because it does not tell you anything.



3.3.1 Exercises
1)What’s gone wrong with this code? Why are the points not blue?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```
The reason is because it's not outside aes().

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

2) Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?
```{r}
?mpg
```
Variables that are categorical are manufacturer, model, trans, drv, fl, and class.
Variable that are continuous are displ, year, cyl, cty, and hwy.
No clue. I just googled.

3) Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = cyl, 
  size = hwy, shape = fl))
```

4) What happens if you map the same variable to multiple aesthetics?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = cyl, size = cyl))
```

5) What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)
```{r}
?geom_point
vignette("ggplot2-specs")
```
Stroke seems to have something to do with a line size? All of them from the looks of it.

6) What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)? Note, you’ll also need to specify x and y.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, colour = displ < 5))
```
The aesthetic is applied to what you used.



3.5.1 Exercises
1) What happens if you facet on a continuous variable?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cyl, y = hwy)) +
  facet_wrap(~ displ)
```

2) What do the empty cells in plot with facet_grid(drv ~ cyl) mean? How do they relate to this plot?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))
```
The facets are empty.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ cyl)
```

3) What plots does the following code make? What does . do?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```
The . excludes the cloumn for the first graph, just repeats multiple rows, so y is repeated. The . exludes the row for the second graph, just repeats multiple columns, so x is repeated.

4) Take the first faceted plot in this section:
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```
What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?
The advantage with facet is that you can look at them each individully side by side. The advantage of colour is you can look at everything as one chart without being confused because it's all differently color coated. With more data, you might want just one graph with everything instead of many individual ones.

5) Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn’t facet_grid() have nrow and ncol arguments?
```{r}
?facet_wrap
```
nrow & ncol are the number of rows and columns. scales for one dimension and shrink? Facet_grid() doesn't have it because it has rows, col which come from vars().

6) When using facet_grid() you should usually put the variable with more unique levels in the columns. Why?
You should because as you put more levels for the row and none for y then the y axis would shrink and be hard to read.



3.6.1 Exercises
1) What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?
```{r}
ggplot(data = mpg) +
  geom_line(mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(x = displ, y = hwy))
```

2) Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.
Maybe a simple single line graph with scatter plots.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```
More colorful and more lines than I thought.

3) What does show.legend = FALSE do? What happens if you remove it?
Why do you think I used it earlier in the chapter?
I don't know.

4) What does the se argument to geom_smooth() do?
Show the area around the line

5) Will these two graphs look different? Why/why not?
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```
No because the layers take the configuration from the ggplot.

6) Recreate the R code necessary to generate the following graphs.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() +
  geom_smooth(se = FALSE)
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() +
  geom_smooth(mapping = aes(group = drv), se = FALSE)
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color=drv)) + 
  geom_point() +
  geom_smooth(se = FALSE)
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color=drv)) +
  geom_smooth(se = FALSE)
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color=drv)) +
  geom_smooth(se = FALSE, mapping = aes(linetype = drv))
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color=drv)) + 
  geom_point(shape = 21, color = "white", stroke = 1)
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color=drv)) +
  geom_point(shape = 21, color = "white", stroke = 2)
```

3.7.1 Exercises
1) What is the default geom associated with stat_summary()? How could you rewrite the previous plot to use that geom function instead of the stat function?
```{r}
ggplot(data = diamonds) +
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )
```

```{r}
ggplot(data = diamonds) +
  geom_pointrange(
    mapping = aes(x = cut, y = depth),
    stat = "summary"
  )
```

```{r}
ggplot(data = diamonds) +
  geom_pointrange(
    mapping = aes(x = cut, y = depth),
    stat = "summary",
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )
```

2) What does geom_col() do? How is it different to geom_bar()?
geom)col() needs x and y to be able to do bar height. They have different defaults.

3) Most geoms and stats come in pairs that are almost always used in concert. Read through the documentation and make a list of all the pairs. What do they have in common?
Some have names in common.

4) What variables does stat_smooth() compute? What parameters control its behaviour?
stat_smooth() computes predicted value, lower and upper value of the confidence interval and standard error.
The parameters that control its behavior are method, formula.

5) In our proportion bar chart, we need to set group = 1. Why? In other words what is the problem with these two graphs?
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop..))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color, y = ..prop..))
```
The problem is the bars make it look like they are all equal.

3.8.1 Exercises
1) What is the problem with this plot? How could you improve it?
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point()
```
It could be that there are too many. The scatter plot already makes it obvious it's increasing. If you condense it, it makes it look nicer. It makes the location of observers easier to figure out.
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point(position = "jitter")
```

2) What parameters to geom_jitter() control the amount of jittering?
To arguments control jitter. Which are width and height. Width controls the amount of vertical displacement and height controls the amount of horizontal displacement.

3) Compare and contrast geom_jitter() with geom_count().
geom_jitter() adds random variation to the location points of the graph. geom_count() sizes the points relative to the number of observations.

4) What’s the default position adjustment for geom_boxplot()? Create a visualisation of the mpg dataset that demonstrates it.
the default position for it is dodge2.
```{r}
ggplot(data = mpg, aes(x = drv, y = hwy, colour = class)) +
  geom_boxplot()
```

3.9.1 Exercises
1) Turn a stacked bar chart into a pie chart using coord_polar().
```{r}
ggplot(mpg, aes(x = factor(1), fill = drv)) +
  geom_bar()
```

```{r}
ggplot(mpg, aes(x = factor(1), fill = drv)) +
  geom_bar(width = 1) +
  coord_polar(theta = "y")
```

2) What does labs() do? Read the documentation.
labs() adds axis titles, plot titles, and a caption to the plot.

3) What’s the difference between coord_quickmap() and coord_map()?
coord_quickmap() function uses an approximate but a faster map. coord_map() uses the Mercator projection.

4) What does the plot below tell you about the relationship between city and highway mpg? Why is coord_fixed() important? What does geom_abline() do?
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() +
  coord_fixed()
```
coord_fixed() ensures that the line produced by geom_abline() is at a 45 degree angle.
